Multi-robot systems play a crucial role in various real-world applications, including autonomous exploration, search and rescue, and logistics. Traditional centralized path planning approaches require a complete map of the environment, which is often unavailable in dynamic and unknown scenarios. To address this limitation, this research proposes a \textbf{decentralized multi-robot path planning framework} using \textbf{Graph Neural Networks (GNNs)}, enabling robots to make autonomous navigation decisions based on local information from neighboring agents.

To enhance coordination and decision-making, we introduce an \textbf{Attention-Augmented GNN}, which improves information aggregation by prioritizing relevant neighbors. This allows for more efficient path planning and obstacle avoidance, even in complex and dynamic environments. The proposed framework is further extended to \textbf{3D navigation}, enabling its applicability to aerial and underwater robotic systems. Additionally, dynamic obstacle handling mechanisms are incorporated to allow real-time adaptation to unpredictable environmental changes.

The effectiveness of the proposed approach is evaluated through extensive simulations, comparing the performance of the \textbf{Attention-Augmented GNN} against a baseline GNN model. Metrics such as \textbf{path efficiency, computational performance, and collision avoidance success rate} are analyzed to demonstrate the advantages of the enhanced model.

By developing a robust decentralized path planning system, this research aims to improve the scalability, adaptability, and real-world feasibility of multi-robot coordination, paving the way for more autonomous and intelligent robotic swarms in uncertain and dynamic environments.

 