I express my sincere gratitude to Dr. Sourav Bhowmick, my advisor and guide, for his unwavering support, insightful guidance, and continuous encouragement throughout this project. His expertise and valuable feedback have been instrumental in shaping my work, and I am deeply appreciative of the time and effort he dedicated to mentoring me.

I also extend my heartfelt thanks to Dr. Deepak Mishra, my co-guide, for his invaluable insights and thoughtful discussions, which significantly contributed to the successful completion of this work. His patience in addressing my doubts and his constructive suggestions have been truly enriching.

Furthermore, I am grateful to the Department of Avionics and the Indian Institute of Space Science and Technology, Thiruvananthapuram, for providing me with the necessary resources and an intellectually stimulating environment to carry out this research. Their support has played a crucial role in the successful execution of this project.

% Lastly, I wish to express my appreciation to my peers, colleagues, friends, and family who have been a source of encouragement and support throughout this journey. Their motivation and assistance have been invaluable in making this project a reality.

%  I would like to extend thanks and appreciation to Dr. Deepak Mishra, Professor $\&$ Head; Dr. Manoj B.S., Professor and Dr. Sheeba Rani, Associate Professor , Department of Avionics, Indian Institute of Space science and Technology for reviewing my presentation and report. \ldots

\vspace{15mm}
\hfill \Author \hspace{5mm}

\hfill {College ID : SC21B127} \hspace{3mm}

%\hfill {URSC ID : T230204} \hspace{4mm}
